---
title: "Engagement with Wikipedia Portal"
author: "Mikhail Popov"
date: "18 July 2016"
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: header.tex
  html_document: default
geometry: margin=0.5in
fontsize: 10pt
---

## Introduction

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
options(digits = 3)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

```{r packages, warning = FALSE, message = FALSE}
library(magrittr)
library(tidyr)
import::from(dplyr, group_by, summarize, keep_where = filter,
             arrange, ungroup, select, rename, mutate, tally)
library(ggplot2)
library(cowplot)
import::from(ggthemes, tufte = theme_tufte)
```

```{r functions, echo = FALSE}
to_ordinal <- function(x) {
  return(vapply(x, toOrdinal::toOrdinal, "VALUE"))
}
most_common <- function(x, ...) {
  if (all(is.na(x))) {
    return(as.character(NA))
  }
  y <- sort(table(x, ...), decreasing = TRUE)
  return(names(y)[1])
}
paste_ <- function(x, ...) {
  if (all(is.na(x))) {
    return(as.character(NA))
  } else {
    paste(x[!is.na(x)], ...)
  }
}
```

```{r read_data}
dashboard <- readr::read_rds("data/dashboard-breakdown-data.rds")
events <- readr::read_rds("data/portal-T139109-data.rds")
```

## Exploratory Data Analysis

```{r dashboard}
dashboard %>%
  keep_where(section_used != "language search") %>%
  ggplot(aes(x = date, y = proportion, color = section_used)) +
  geom_line() +
  scale_y_continuous("Proportion", labels = scales::percent_format(),
                     breaks = seq(0, 1, 0.1)) +
  geom_vline(xintercept = as.numeric(lubridate::ymd("2016-06-17")), linetype = "dashed") +
  tufte(base_size = 12) +
  labs(title = "Portal Action Breakdown", subtitle = "Aggregate data downloaded from datasets.wikimedia.org")
```

```{r}
events %>%
  group_by(date, section_used = ifelse(is.na(section_used), "no action", section_used)) %>%
  tally() %>%
  mutate(proportion = n/sum(n)) %>%
  ggplot(aes(x = date, y = proportion, color = section_used)) +
  geom_line() +
  scale_y_continuous("Proportion", labels = scales::percent_format(),
                     breaks = seq(0, 1, 0.1)) +
  geom_vline(xintercept = as.numeric(lubridate::ymd("2016-06-17")), linetype = "dashed") +
  tufte(base_size = 12) +
  labs(title = "Portal Action Breakdown", subtitle = "No filtering")
```

The difference is actually caused by [a line deletion](https://github.com/wikimedia/wikimedia-discovery-golden/commit/c3cbe13f9cc29e9ef92c8d79b2c545d1afc238a7#diff-c55f3a392650f4257983a2ee3bf112cc) from the script that processes the data fetched from the EL database. Here's the same data but with a filtering that the script previously had.

```{r}
p1 <- events %>%
  arrange(type) %>%
  dplyr::distinct(session_id, .keep_all = TRUE) %>%
  group_by(date, section_used = ifelse(is.na(section_used), "no action", section_used)) %>%
  tally() %>%
  mutate(proportion = n/sum(n)) %>%
  ggplot(aes(x = date, y = proportion, color = section_used)) +
  geom_line() +
  scale_y_continuous("Proportion", labels = scales::percent_format(),
                     breaks = seq(0, 1, 0.1)) +
  geom_vline(xintercept = as.numeric(lubridate::ymd("2016-06-17")), linetype = "dashed") +
  tufte(base_size = 12) +
  ggtitle("Portal Action Breakdown", subtitle = "With 'last action taken' filtering") +
  theme(legend.position = "bottom")
print(p1)
```

The reasoning behind that change was that this panel is supposed to be counts/proportions of ALL events, even if one user had a single session with hundreds or thousands of events. Previously, the breakdown referred to the last action in each session. A separate dashboard panel was created for "actions on first visit". This change has been [reverted](https://gerrit.wikimedia.org/r/#/c/299085/), the data removed, and backfilled on 14 July 2016.

## Why unfiltered "no action" numbers are so big

```{r}
clickthroughs <- data.table::as.data.table(events)[, {
  data.frame(clickthrough = any(type == "clickthrough"),
             clicks = sum(type == "clickthrough"),
             most_common_section = most_common(section_used),
             clicks_to_that_section = sum(section_used == most_common(section_used), na.rm = TRUE),
             date = as.Date(min(ts)),
             stringsAsFactors = FALSE)
}, by = c("identity", "country_name", "visit")]
```

\begin{landscape}

\subsection*{Top sessions by number of clicks per session}

```{r, echo = FALSE}
top_ids_by_clicks <- clickthroughs[, list(ctr = max(clicks)), by = "identity"] %>%
  arrange(desc(ctr)) %>%
  head(10) %>%
  { .$identity }
```

```{r, results = 'asis'}
events %>%
  keep_where(identity %in% top_ids_by_clicks) %>%
  group_by(identity, country = country_name, device, os, browser, visit) %>%
  summarize(clickthroughs = sum(type == "clickthrough"),
            `sections used` = length(unique(section_used)) - 1,
            `most used section` = most_common(section_used),
            `clicks on that section` = sum(section_used == `most used section`, na.rm = TRUE)) %>%
  ungroup %>%
  arrange(desc(clickthroughs)) %>%
  head(10) %>%
  mutate(visit = to_ordinal(visit)) %>%
  select(-identity) %>%
  knitr::kable(caption = "Counts of clicks per visit.", format = "latex")
```

\subsection*{Top sessions by number of visits per session}

```{r, echo = FALSE}
top_ids_by_visit <- clickthroughs[, list(visits = .N), by = "identity"] %>%
  arrange(desc(visits)) %>%
  head(10) %>%
  { .$identity }
```

```{r, results = 'asis'}
events %>%
  keep_where(identity %in% top_ids_by_visit) %>%
  group_by(identity, country = country_name, device, os, browser, visit) %>%
  summarize(clickthroughs = sum(type == "clickthrough"),
            `sections used` = length(unique(section_used)) - 1,
            `most used section` = most_common(section_used),
            `clicks on that section` = sum(section_used == `most used section`, na.rm = TRUE),
            first_ts = min(ts), last_ts = max(ts)) %>%
  group_by(identity) %>%
  summarize(visits = n(),
            `session length` = tolower(lubridate::seconds_to_period(as.numeric(difftime(max(last_ts), min(first_ts), units = "secs")))),
            country = country[1], device = device[1], os = os[1], browser = browser[1],
            `total sections used` = sum(`sections used`),
            `most used sections` = paste_(unique(`most used section`), collapse = ", "),
            `most used section` = (`most used section`[order(`clicks on that section`, decreasing = TRUE)])[1],
            `total clicks` = sum(clickthroughs),
            `max clickthroughs per visit` = max(clickthroughs),
            `median clickthroughs per visit` = median(clickthroughs),
            `clickthrough rate` = sum(clickthroughs > 0)/visits) %>%
  arrange(desc(visits)) %>%
  select(-identity) %>%
  knitr::kable(caption = "Top 10 sessions by number of visits from 13 May 2016 to 13 July 2016", format = "latex")
```

\end{landscape}

## How engagement varies by country

```{r}
clickthroughs[!is.na(clickthroughs$country_name),
     list(sessions = length(unique(identity)),
          visits = .N,
          `visits with clickthrough` = sum(clickthrough)/.N),
     by = c("date", "country_name")] %>%
  group_by(date) %>%
  mutate(`total daily visits` = visits/sum(visits),
         `total daily sessions` = sessions/sum(sessions)) %>%
  dplyr::top_n(5, visits) %>%
  select(-c(visits, sessions)) %>%
  ungroup %>%
  gather(`proportion of`, proportion,
         c(`total daily visits`, `total daily sessions`, `visits with clickthrough`)) %>%
  ggplot(aes(x = date, y = proportion, color = country_name)) +
  geom_line() +
  scale_y_continuous(labels = scales::percent_format(),
                     breaks = seq(0, 1, 0.1)) +
  facet_wrap(~ `proportion of`, nrow = 3, scales = "free_y") +
  tufte(base_size = 12) +
  ggtitle("Top 10 countries by number of visits",
          subtitle = "From 13 May 2016 to 13 July 2016")
```

## Potential metrics as alternatives or in addition to current ones

### Most common section used in a visit

```{r, results = 'asis'}
p2 <- clickthroughs[, list(visits = .N,
                     sessions = length(unique(identity))),
              by = c("date", "most_common_section")] %>%
  group_by(date) %>%
  mutate(`Proportion (Sessions)` = sessions/sum(sessions),
         most_common_section = ifelse(is.na(most_common_section), "none", most_common_section)) %>%
  rename(`Most common section used in a visit` = most_common_section,
         `Total visits` = visits, `Total sessions` = sessions) %>%
  ggplot(aes(x = date, y = `Proportion (Sessions)`,
             color = `Most common section used in a visit`)) +
  geom_line() +
  scale_y_continuous(labels = scales::percent_format(),
                     breaks = seq(0, 1, 0.1)) +
  tufte(base_size = 12) +
  ggtitle("Most commonly clicked sections per visits") +
  theme(legend.position = "bottom")
print(p2)
```

## Most common section clicked on across all visits in a session

```{r, results = 'asis'}
p3 <- clickthroughs %>%
  group_by(identity, most_common_section) %>%
  summarize(clicks_to_that_section = sum(clicks_to_that_section),
            date = min(date)) %>%
  group_by(date, identity) %>%
  dplyr::top_n(1, clicks_to_that_section) %>%
  group_by(date) %>%
  mutate(most_common_section = ifelse(is.na(most_common_section), "none", most_common_section)) %>%
  group_by(date, `Most commonly clicked section in session` = most_common_section) %>%
  summarize(Sessions = n()) %>%
  mutate(`Proportion (Sessions)` = Sessions/sum(Sessions)) %>%
  ggplot(aes(x = date, y = `Proportion (Sessions)`,
             color = `Most commonly clicked section in session`)) +
  geom_line() +
  scale_y_continuous(labels = scales::percent_format(),
                     breaks = seq(0, 1, 0.1)) +
  tufte(base_size = 12) +
  ggtitle("Number of sessions where each particular section was the most clicked section in each user's session across all their visits",
       subtitle = "From 13 May 2016 to 13 July 2016") +
  theme(legend.position = "bottom")
print(p3)
```
