---
title: "Non-Search Engine Referred Traffic to Wikipedia.org"
author: "Mikhail Popov"
date: "March 16, 2016"
output: pdf_document
fontsize: 12pt
---
\renewcommand{\abstractname}{Executive Summary}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(magrittr)
library(dplyr)
library(ggplot2)
top_n2 <- function(x, n, wt, var, f = sum) {
  expr <- paste0("dplyr::top_n(x, ", n, ", ", deparse(substitute(wt)), ")")
  y <- eval(parse(text = expr))
  `%notin%` <- function(x, y) {
    return(!(x %in% y))
  }
  filter_dots <- eval(parse(text = paste0("list(~", deparse(substitute(var)), "%notin% y$", deparse(substitute(var)), ")")))
  summary_dots <- eval(parse(text = paste0("list(~paste0('...', length(", deparse(substitute(var)),
                                           "), ' others...'), ~f(", deparse(substitute(wt)), "))")))
  z <- x %>%
    filter_(.dots = filter_dots) %>%
    summarize_(.dots = summary_dots)
  names(z) <- c(deparse(substitute(var)), deparse(substitute(wt)))
  return(rbind(y, z))
}
```

\begin{abstract}

On 10 March 2016 we deployed a change to the External Traffic part of the Portal dashboard that broke down the "non-search engine" traffic into direct traffic and "referred by something other than search engine" traffic. Approximately 10\% of all traffic to wikipedia.org is referred by something other than a search engine.

In this report we dove deeper into that 10\% traffic to see where the Portal users are coming from:
\begin{itemize}
  \item About 77\% of referred traffic (which accounts for 10\% of overall traffic to wikipedia.org) is from Wikimedia projects, most notably wikipedia.org itself, which accounts for nearly all of the internally-referred traffic.
  \item External traffic (21.6\% of referred traffic) is largely accounted by Lightning newtab (a Chrome extension, 23\% of external traffic) and AlexaSurfing (Alexa rank booster, 10.3\% of external traffic).
\end{itemize}
\end{abstract}

```{r, fig.width = 6, fig.height = 4, fig.align = 'center'}
summary_traffic_data <- readr::read_csv("traffic_summary.csv")
ggplot(data = summary_traffic_data, aes(x = date, y = percentage, color = traffic)) +
  geom_line() +
  scale_color_discrete("Traffic Source") +
  scale_y_continuous(breaks = seq(0, 1, 0.1),
                     labels = scales::percent_format()) +
  labs(title = "Sources of Daily Traffic to Wikipedia.org",
       x = "Date", y = "Percentage of Total Traffic") +
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "vertical")
```

\newpage

## Analysis of 10% Traffic

This report looks at web request logs from 8 March 2016 to 15 March 2016, filtered by presence of a referer but specifically not one of the major search engines the refinery recognizes (Baidu, Bing, Google, Yahoo!, Yandex). External traffic accounts for 21.6%, internal for 76.8%, and unknown referrers for 0.5% on average (median).

|Date       |External |Internal |Unknown |
|:----------|:--------|:--------|:-------|
|2016-03-08 |17.86%   |81.82%   |0.32%   |
|2016-03-09 |16.41%   |75.49%   |8.11%   |
|2016-03-10 |18.67%   |80.83%   |0.50%   |
|2016-03-11 |23.03%   |76.52%   |0.44%   |
|2016-03-12 |22.65%   |77.08%   |0.28%   |
|2016-03-13 |20.56%   |78.93%   |0.51%   |
|2016-03-14 |24.18%   |75.19%   |0.63%   |
|2016-03-15 |24.37%   |74.76%   |0.88%   |

We explored some of the unknown and internal referrers, and found the following statistics:

|                                                    | Requests| % of 10% Traffic|
|:---------------------------------------------------|--------:|----------------:|
|Referred by ourselves (wikipedia.org)               |  7709525|           76.84%|
|Unknown referrer format                             |   175256|            1.75%|
|Internal referer (non-Portal)                       |    96253|            0.96%|
|Referred by Wikimedia Nederland (wikimedia.nl)      |    54600|            0.54%|
|Empty scheme                                        |    31054|            0.31%|
|Referred from a file                                |      119|          <0.002%|
|Referred by hndUnblock.cgi on Cisco/Linksys routers |        7|         <0.0001%|
|Referred by IPv6                                    |        6|         <0.0001%|
|Garbage referer metadata                            |        3|         <0.0001%|
|Referred by mailto link                             |        2|         <0.0001%|

For the analysis of external traffic, we excluded the requests matching the categories above.

\newpage

### Analysis of Internal Traffic

The table below shows the top 11 WMF projects with respect to percentage of internally-referred traffic they accounted for.

|Domain                  |% of Internal Traffic |
|:-----------------------|:---------------------|
|wikipedia.org           |98.766%               |
|de.wikipedia.org        |0.811%                |
|wikivoyage.org          |0.083%                |
|wikimedia.org           |0.064%                |
|en.m.wikipedia.org      |0.048%                |
|es.m.wikipedia.org      |0.036%                |
|es.wikipedia.org        |0.032%                |
|en.wikipedia.org        |0.027%                |
|wikimediafoundation.org |0.018%                |
|wikinews.org            |0.018%                |
|wiktionary.org          |0.015%                |
|...202 others...        |0.081%                |

### Analysis of External Traffic
```{r}
external_traffic_data <- readr::read_csv("referrals.csv.gz", col_types = "ccccccc")
external_traffic_data$date %<>% lubridate::ymd()
```

The two major sources of traffic to Wikipedia.org appear to be...Lighting newtab (a Chrome extension) and AlexaSurfing: Alexa Rank Booster (an Alexa traffic exchange service).

```{r}
external_traffic_data %>%
  group_by(domain) %>%
  summarize(n = n()) %>%
  mutate(n = n/sum(n)) %>%
  arrange(desc(n)) %>%
  top_n2(10, n, domain) %>%
  mutate(n = sprintf("%.2f%%", 100*n)) %>%
  set_colnames(c("Source", "% of External Traffic")) %>%
  knitr::kable(format = "markdown")
```

\newpage

```{r, fig.width = 7, fig.height = 8, fig.align = 'center'}
external_traffic_data %>%
  group_by(date, domain) %>%
  summarize(n = n()) %>%
  mutate(n = n/sum(n)) %>%
  top_n(5, n) %>%
  ggplot(data = ., aes(x = date, y = n, color = domain)) +
  geom_line(size = 1.25) +
  scale_y_continuous("% of External Traffic", labels = scales::percent_format()) +
  scale_color_brewer(name = "Referrer", type = "qual", palette = "Dark2") +
  labs(x = "Date", title = "Top 5 Sources of External Traffic to Wikipedia.org") +
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "vertical")
```

There is some fluctuation in traffic from those sources, but we can still get a good picture of the top referral traffic in a week.
